---
title: "PFI Exploratory"
author: "Evan “Pete” Walsh"
date: "October 15, 2014"
output: html_document
---

>Load data and libraries.

```{r message=FALSE}
pfi <- read.csv("/Users/marianwaitwalsh/Classes and Work/PFI/PFI_clean.csv")
library(dplyr)
library(ggplot2)
```

>Variables are grouped into five different categories, denoted by ```item_type```. These categories are ```Cost```, ```Crop```, ```Expense```, ```Income```, and ```Labor```.

```{r}
unique(pfi$item_type)
```


>Create a numeric version of ```value```.

```{r warning=FALSE}
pfi$value2 <- as.numeric(as.character(pfi$value))
```

>Here is a summary of the ```crop``` type produced by each field.

```{r}
crop <- pfi %>%
  filter(item_type=="Crop") %>%
  group_by(field_id) %>%
  summarize(crop_types = unique(value))
crop
```

>Let's make a new column in ```pfi``` that just has the crop type by field.

```{r}
pfi$crop <- NA
pfi$crop[which(pfi$field_id == "Boone1")] <- "Corn"
pfi$crop[which(pfi$field_id == "Boone2")] <- "SB"
pfi$crop[which(pfi$field_id == "Thompson1")] <- "Corn"
pfi$crop[which(pfi$field_id == "Thompson2")] <- "SB"
pfi$crop[which(pfi$field_id == "Thompson3")] <- "Corn"
pfi$crop[which(pfi$field_id == "Thompson4")] <- "Oats"
pfi$crop[which(pfi$field_id == "Thompson5")] <- "Hay"
pfi$crop <- as.factor(pfi$crop)
```

>We will now examine the category ```Income```.

```{r}
inc <- pfi %>%
  filter(item_type=="Income") %>%
  group_by(field_id, item) %>%
  summarize(mean = mean(value2),
            s = sd(value2),
            n = length(value2))
inc
inc2 <- pfi %>%
  select(field_id, year, item, crop, value2) %>%
  filter(item=="Crop_Income") %>%
  group_by(field_id) %>%
  arrange(year)
qplot(year, value2, data=inc2, colour=crop, facets=~field_id) +
  ylab("Crop Income") + stat_smooth(method=loess)
inc3 <- pfi %>%
  select(field_id, year, item, crop, value2) %>%
  filter(item=="Yield_Per_Acre_Bu_per_pound", field_id != "Thompson5") %>%
  group_by(field_id) %>%
  arrange(year)
qplot(year, value2, data=inc3, colour=crop, facets=~field_id) + 
  ylab("Yield / acre bu / #") + stat_smooth(method=loess)
```

>Questions at this point:
<ol><li>Why is the Yield_Per_Acre_Bu / # so high for Thompson5?</li>
<li>Which variable should we be most interested in?</li></ol>

>Here is a summary of the only ```Cost``` variable, which is ```Cost_per_bu_Or_pound```.

```{r}
cost <- pfi %>%
  filter(item_type=="Cost") %>%
  group_by(field_id, item) %>%
  summarize(mean = mean(value2),
            s = sd(value2),
            n = length(value2))
cost
cost2 <- pfi %>%
  select(field_id, year, item, crop, value2) %>%
  filter(item=="Cost_per_bu_Or_pound", field_id != "Thompson5") %>%
  group_by(field_id) %>%
  arrange(year)
qplot(year, value2, data=cost2, colour = crop, facets=~field_id) + 
  ylab("Cost / bu or #") + stat_smooth(method=loess)
```

>Let's look at the ```labor``` category now.

```{r}
labor <- pfi %>%
  filter(item_type=="Labor") %>%
  group_by(field_id, item) %>%
  summarize(mean = mean(value2),
            s = sd(value2),
            n = length(value2))
labor2 <- pfi %>%
  select(field_id, year, item, crop, value2) %>%
  filter(item == "Labor_Return") %>%
  group_by(field_id) %>%
  arrange(year)
labor3 <- pfi %>%
  select(field_id, year, item, crop, value2) %>%
  filter(item == "LaborandMR_dollar_per_A") %>%
  group_by(field_id) %>%
  arrange(year)
qplot(year, value2, data=labor2, colour=crop, facets=~field_id) + 
  ylab("Labor return")
qplot(year, value2, data=labor3, colour=crop, facets=~field_id) + 
  ylab("Labor MR $ / Acre") + stat_smooth(method=loess)
```

