---
title: "Sources of Differences in Income"
author: "Maggie Johnson"
output: html_document
---

```{r, message=FALSE, echo=FALSE, warning=FALSE}
##### Libraries #####
library(knitr)
library(ggplot2)
library(reshape2)
library(plyr)
library(gridExtra)
```

```{r, echo=FALSE, warning=FALSE}
##### Read in Data #####
setwd("C:/Users/majohnso/Documents/GitHub/PFI")
dat <- read.csv("data/PFI_clean.csv")
var <- read.csv("data/var_map.csv")

###### sooo messy, fix later
### Transform the Data ###
dat<-dat[complete.cases(dat),]
dat$field <- factor(paste(dat$farmer, dat$field_id, sep=""))
years <- split(dat, dat$year)
byyear <- lapply(years, function(u) acast(u, field~item, value.var="value"))

years2 <- data.frame(year=rep(unique(dat$year), each=7), field=rep(levels(dat$field), 25), do.call(rbind, byyear))
years2$crop <- as.factor(unique(data.frame(dat$year,dat$field, dat$crop))[,3])
years2 <- years2[complete.cases(years2),]
years2$practice<-as.factor(sub("[0-9].*", "", years2$field))

#### Overall Income 
total_income_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_income = mean(LaborandMR_dollar_per_A))
total_income<-ddply(total_income_crop, .(year, practice), summarize,
              mean_income=mean(mean_income))
```

### Overall Profit 
```{r, fig.pos='h', fig.width=9, fig.align='center'}
theme_set(theme_bw(base_size=12))
p1<-qplot(data=total_income, x=year, y=mean_income, group=practice, colour=practice, geom="line", ylab="Profit", xlab="Year", main="Overall Profit") + theme(legend.position="bottom")
diff_pextra<-subset(total_income, practice=="Thompson")$mean_income - subset(total_income, practice=="Boone")$mean_income
p2<-qplot(x=unique(total_income$year), y=diff_pextra, geom="line", ylab="Difference in Profit", xlab="Year", main="Thompson Advantage") + geom_hline(aes(yintercept=0)) 
grid.arrange(p1,p2,ncol=2)

qplot(data=subset(total_income_crop, crop %in% c("Corn","SB")), x=year, y=mean_income, colour=practice, group=practice, geom="line") + facet_wrap(~crop) + theme(legend.position="bottom")


```

This is the difference in average profit when profit is calculated as the sum of all income (including labor return) minus expenses.


```{r, echo=FALSE, warning=FALSE}
### Proportion of profit due to labor return
years2$Total_Income <- years2$Pasture_per_Stubble + years2$Straw_Income + years2$Residue_Income + years2$Crop_Income
lr_effect_crop<-ddply(years2, .(year, practice, crop), summarize,
                   prop_LR = mean(abs((Labor_Return)/Total_Income)))
lr_effect<-ddply(lr_effect_crop, .(year, practice), summarize,
              prop_LR=mean(prop_LR))

lr_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_LR = mean(Labor_Return))
lr<-ddply(lr_crop, .(year, practice), summarize,
              mean_LR=mean(mean_LR))

```

### Proportion of Income Due to Labor Return
```{r, fig.pos='h', fig.width=9, fig.align='center'}
qplot(data=lr_effect_crop, x=year, y=prop_LR, colour=crop, group=crop, geom="line") + facet_wrap(~practice)

qplot(data=lr_effect, x=year, y=prop_LR, colour=practice, group=practice, geom="line") 
```

Thompson farms have higher average labor returns in general, but this shows that the labor return appears to account for a higher proportion of income in the Thompson farms than in the Boone farms. Do the work yourself! 

### Proportion of Thompson Advantage Due to Labor Return
```{r, fig.pos='h', fig.width=7, fig.align='center'}
diff_profit<-subset(total_income, practice=="Thompson")$mean_income - subset(total_income, practice=="Boone")$mean_income
diff_lr<-subset(lr, practice=="Thompson")$mean_LR - subset(lr, practice=="Boone")$mean_LR

df2<-data.frame(year=unique(total_income$year), diff_profit, diff_lr, prop_lr=diff_lr/diff_profit)
qplot(data=df2, x=year, y=prop_lr, geom="line", xlab="Year", ylab="Proportion of Profit Difference") 
```

Consistently, between 10-20% of Thompson's advantage is due to the the difference in labor return between Boone and Thompson farms.
