```{r}
#dat %>%
#  filter(item_type == "Revenue")

```

```{r, echo=FALSE, warning=FALSE}

###### sooo messy, fix later
### Transform the Data ###
dat2<-dat[complete.cases(dat),]
dat2$field <- factor(paste(dat2$farmer, dat2$field_id, sep=""))
years <- split(dat2, dat2$year)
byyear <- lapply(years, function(u) acast(u, field~item, value.var="value"))

years2 <- data.frame(year=rep(unique(dat2$year), each=7), field=rep(levels(dat2$field), 25), do.call(rbind, byyear))
years2$crop <- as.factor(unique(data.frame(dat2$year,dat2$field, dat2$crop))[,3])
years2 <- years2[complete.cases(years2),]
years2$practice<-as.factor(sub("[0-9].*", "", years2$field))

#### Overall Income 
total_income_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_income = mean(LaborandMR_dollar_per_A))
total_income<-ddply(total_income_crop, .(year, practice), summarize,
              mean_income=mean(mean_income))
```

An overall look at income leads us directly to the question of profit. Figure \ref{fig:profit_diff} demonstrates the difference in overall profit between the Thompson farm and his neighbor. Profit is calculated as the sum of all income (including labor return) minus expenses. It is clear that Thompson's profit is consistently higher than his neighbors. Accordingly, Thompson's income from corn and soy are higher for every year, as can been seen in figure \ref{fig:income_crop}. The question of interest is to determine where this profit is coming from. Better practices, better land, labor return, etc.


```{r, fig.cap="\\label{fig:profit_diff}The difference in overall profit between the Thompson farm and his neighbor. Profit is calculated as the sum of all income (including labor return) minus expenses. The left plot is the overall profit, while the right plot shows the differences between Thompson and his neighbor."}
p1<-qplot(data=total_income, x=year, y=mean_income, group=practice, colour=practice, geom="line", ylab="Profit", xlab="Year") + theme(legend.position="bottom")
diff_pextra<-subset(total_income, practice=="Thompson")$mean_income - subset(total_income, practice=="Boone")$mean_income
p2<-qplot(x=unique(total_income$year), y=diff_pextra, geom="line", ylab="Difference in Profit (Thompson - Boone)", xlab="Year") + geom_hline(aes(yintercept=0)) 
grid.arrange(p1,p2,ncol=2)


```

```{r, fig.cap="\\label{fig:income_crop}Average income by crop for corn and soy over time compared between Thompson and Boone county. Thompson's income is higher for every year for both crops."}
qplot(data=subset(total_income_crop, crop %in% c("Corn","SB")), x=year, y=mean_income, colour=practice, group=practice, geom="line") + facet_wrap(~crop) + theme(legend.position="bottom")


```

## Effect of Labor Return 
Here we explore the effect of Labor Return on overall profit between the Thompson and Boone practices. From what we understand, Labor Return is the amount of money Thompson paid himself for the work he did on his own, instead of hiring out. We explore how much of Thompson's advantage is due to Labor Return.


```{r, echo=FALSE, warning=FALSE}
### Proportion of profit due to labor return
years2$Total_Income <- years2$Pasture_per_Stubble + years2$Straw_Income + years2$Residue_Income + years2$Crop_Income
lr_effect_crop<-ddply(years2, .(year, practice, crop), summarize,
                   prop_LR = mean(abs((Labor_Return)/Total_Income)))
lr_effect<-ddply(lr_effect_crop, .(year, practice), summarize,
              prop_LR=mean(prop_LR))

lr_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_LR = mean(Labor_Return))
lr<-ddply(lr_crop, .(year, practice), summarize,
              mean_LR=mean(mean_LR))

```


```{r, fig.cap="\\label{fig:LR_prop}The proportion of Thompson's and Boone county total income that is accounted for by Labor return. The right plot shows the same, but split by crop type. Labor return appears to account for a higher proportion of income in the Thompson farms than in the Boone farm."}
qplot(data=lr_effect, x=year, y=prop_LR, colour=practice, group=practice, geom="line") + theme(legend.position="bottom") -> plot1

qplot(data=lr_effect_crop, x=year, y=prop_LR, colour=practice, group=practice, geom="line") + 
  facet_wrap(~crop) + 
  theme(legend.position="bottom") -> plot2

grid.arrange(plot1, plot2, ncol = 2)


```

Figure \ref{fig:LR_prop} shows what proportion of Thompson and Boone total income is accounted for by Labor Return.

Thompson farms have higher average labor returns in general, but this shows that the labor return appears to account for a higher proportion of income in the Thompson farms than in the Boone farms. Do the work yourself, if you can! 

```{r, fig.cap="\\label{fig:LR_profit}The proportion of profit difference (Thompson - Boone) that is accounted by labor returns. Once again, this is showing labor returns to be a major contributor to Thompson's profit advantage over the Boone County neighbor.", fig.height = 3, fig.width=9}
diff_profit<-subset(total_income, practice=="Thompson")$mean_income - subset(total_income, practice=="Boone")$mean_income
diff_lr<-subset(lr, practice=="Thompson")$mean_LR - subset(lr, practice=="Boone")$mean_LR

df2<-data.frame(year=unique(total_income$year), diff_profit, diff_lr, prop_lr=diff_lr/diff_profit)
qplot(data=df2, x=year, y=prop_lr, geom="line", xlab="Year", ylab="Prop. of Profit Diff.") + ylim(c(0, .5))
```

Consistently, between 10-20% of Thompson's profit advantage is due to the the difference in labor return between Boone and Thompson farms. This is seen in figure \ref{fig:LR_profit}, which shows the proportion of profit difference (Thompson - Boone) that is accounted for by labor returns. Once again, this suggests labor returns are a major contributor to Thompson's profit advantage.

## Effect of Livestock

```{r, echo=FALSE, warning=FALSE}
### Income and Expenses ###
income_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_income = mean(Crop_Income))
income<-ddply(income_crop, .(year, practice), summarize,
              mean_income=mean(mean_income))

expense_crop<-ddply(dat2[dat2$item_type=="Expense",], .(year, field, crop), summarize,
                    expense = sum(value))
expense_crop$practice<-as.factor(sub("[0-9].*", "", expense_crop$field))

### Remove fertilizer
# expense_nofert_crop<-ddply(dat2[dat2$item_type=="Expense" & dat2$item!="Purch_Pert" & dat2$item!="Maunure_Charge",], .(year, field, crop), summarize,
#                     expense = sum(value))

expense_nofert_crop<-ddply(subset(dat2, item_type=="Expense" & item!="Purch_Pert" & item!="Maunure_Charge" & item!="Windrow_Oats" & item!="Stubble_Costs" & item!="Straw_Costs"), .(year, field, crop), summarize,
                    expense = sum(value))

expense_nofert_crop$practice<-as.factor(sub("[0-9].*", "", expense_crop$field))

expense_crop2<-ddply(expense_crop, .(year, practice, crop), summarize,
                     expense=mean(expense))
expense_nofert_crop2<-ddply(expense_nofert_crop, .(year, practice, crop), summarize,
                     expense_nofert=mean(expense))

expenses<-merge(expense_crop2, expense_nofert_crop2)

expenses_overall<-ddply(expenses, .(year, practice), summarize,
               mean_expense=mean(expense),
               mean_expense_nofert=mean(expense_nofert))

add_income<-ddply(years2, .(year, practice, crop), summarize,
                extra_income=mean(Pasture_per_Stubble + Residue_Income + Straw_Income))

income_crop$extra<-add_income$extra_income
income2<-ddply(income_crop, .(year, practice), summarize,
               mean_allincome=mean(mean_income+extra))

df<-merge(income, income2)
df<-merge(df, expenses_overall)

df$rawprofit <- df$mean_income - df$mean_expense
df$profit_extra <- income2$mean_allincome - df$mean_expense
df$rawprofit_nofert <- df$mean_income - df$mean_expense_nofert

```

The profit and losses from livestock on the Thompson farms are not included in the data. It is suggested that Thompson fertilized his fields using manure from his livestock, and subsequently has an expense due to spreading manure, but no expenses for purchasing fertilizer. The Boone expenses on the other hand include a large sum (relative to other expenses) for purchasing fertilizer. As livestock costs and profits are not included, it seems reasonable to explore excluding these variables from the analysis. Note: this analysis was done without additional income and expenses from straw, stubble, and residue in order to further even the comparison since Thompson's neighbor did not receive income from these items. (The variables excluded are \texttt{Purch\_Pert}, \texttt{Manure\_Charge}, \texttt{Pasture\_per\_Stubble}, \texttt{Straw\_Income}, \texttt{Residue\_Income}, \texttt{Windrow\_Oats}, \texttt{Stubble\_Costs}, and \texttt{Straw\_Costs}).

```{r, fig.cap="\\label{fig:fert_exp}Total expenses with/without fertilizer and additional practices between Thompson and Boone."}
df %>%
  gather(fertilizer, mean_expense, mean_expense, mean_expense_nofert) %>% 
  mutate(fertilizer = ifelse(fertilizer == "mean_expense", "With fertilizer", "W/o fertilizer")) %>%
  ggplot() +
  geom_line(aes(year, mean_expense, colour = practice)) + 
  facet_wrap(~fertilizer) +
  theme(legend.position="bottom") +
  ylab("Expense/Acre")

```

Figure \ref{fig:fert_exp} shows that removing fertilizer and additional practices (stubble, residue, etc.) lessens the difference in overall expenses between Thompson and Boone practices, but Thompson still maintained consistently less expenses than his neighbor across the years.

```{r, fig.cap="\\label{fig:exp_corn}Expenses for corn and soybeans with and without fertilizer costs."}
expenses %>%
  filter(crop %in% c("Corn", "SB")) %>%
  gather(fertilizer, expense, expense, expense_nofert) %>% 
  mutate(fertilizer = ifelse(fertilizer == "expense", "With fertilizer", "W/o fertilizer")) %>%
  ggplot() +
  geom_line(aes(year, expense, colour = practice)) + 
  facet_grid(fertilizer ~ crop) +
  theme(legend.position="bottom") +
  ylab("Expense/Acre")

```

When we look at the effect of fertilizer on the difference in expenses for just the corn and soybean crops in figure \ref{fig:exp_corn}, we see that the additional expenses attributed to corn are nearly the same for Boone and Thompson between 1995 - 2005. Additionally, figure \ref{fig:prof_corn} shows that the difference in overall profit due to corn between Boone and Thompson is less when we remove the effect of livestock.

```{r, fig.cap="\\label{fig:prof_corn}Profit from corn with and without fertilizer and additional practices expenses."}
df %>%
  gather(fertilizer, rawprofit, rawprofit, rawprofit_nofert) %>% 
  mutate(fertilizer = ifelse(fertilizer == "rawprofit", "With fertilizer", "W/o fertilizer")) %>%
  ggplot() +
  geom_line(aes(year, rawprofit, colour = practice)) + 
  facet_wrap(~fertilizer) +
  theme(legend.position="bottom") +
  ylab("Profit/Acre")

```

Removal of livestock effects seems to lessen the difference in expenses and profit between the Boone and Thompson averages, but the Thompson farm still consistently comes out ahead. This suggests that while Thompson's utilization of his livestock is not the sole source of his advantage, it plays a role in increasing his profits over the Boone neighbors. 
