## Effect or Labor Return 
Here we explore the effect of Labor Return on overall profit between the Thompson and Boone practices. From what we understand, Labor Return is the amount of money Thompson paid himself for the work he was doing on his own, instead of hiring out. We explore how much of Thompson's advantage is due to Labor Return.

```{r, echo=FALSE, warning=FALSE}

###### sooo messy, fix later
### Transform the Data ###
dat2<-dat[complete.cases(dat),]
dat2$field <- factor(paste(dat2$farmer, dat2$field_id, sep=""))
years <- split(dat2, dat2$year)
byyear <- lapply(years, function(u) acast(u, field~item, value.var="value"))

years2 <- data.frame(year=rep(unique(dat2$year), each=7), field=rep(levels(dat2$field), 25), do.call(rbind, byyear))
years2$crop <- as.factor(unique(data.frame(dat2$year,dat2$field, dat2$crop))[,3])
years2 <- years2[complete.cases(years2),]
years2$practice<-as.factor(sub("[0-9].*", "", years2$field))

#### Overall Income 
total_income_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_income = mean(LaborandMR_dollar_per_A))
total_income<-ddply(total_income_crop, .(year, practice), summarize,
              mean_income=mean(mean_income))
```

### Overall Profit 
```{r, fig.pos='H', fig.width=9, fig.align='center', echo=FALSE}
theme_set(theme_bw(base_size=12))
p1<-qplot(data=total_income, x=year, y=mean_income, group=practice, colour=practice, geom="line", ylab="Profit", xlab="Year", main="Overall Profit") + theme(legend.position="bottom")
diff_pextra<-subset(total_income, practice=="Thompson")$mean_income - subset(total_income, practice=="Boone")$mean_income
p2<-qplot(x=unique(total_income$year), y=diff_pextra, geom="line", ylab="Difference in Profit", xlab="Year", main="Thompson Advantage") + geom_hline(aes(yintercept=0)) 
grid.arrange(p1,p2,ncol=2)

qplot(data=subset(total_income_crop, crop %in% c("Corn","SB")), x=year, y=mean_income, colour=practice, group=practice, geom="line") + facet_wrap(~crop) + theme(legend.position="bottom")


```

These plots demonstrate the difference in overall profit between the Thompson farm and Boone averages. Profit is calculated as the sum of all income (including labor return) minus expenses. It is clear that Thompson's profit is consistently higher than the Boone farm averages. The question of interest is to determine where this profit is coming from. Better practices, better land, labor return, etc.

```{r, echo=FALSE, warning=FALSE}
### Proportion of profit due to labor return
years2$Total_Income <- years2$Pasture_per_Stubble + years2$Straw_Income + years2$Residue_Income + years2$Crop_Income
lr_effect_crop<-ddply(years2, .(year, practice, crop), summarize,
                   prop_LR = mean(abs((Labor_Return)/Total_Income)))
lr_effect<-ddply(lr_effect_crop, .(year, practice), summarize,
              prop_LR=mean(prop_LR))

lr_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_LR = mean(Labor_Return))
lr<-ddply(lr_crop, .(year, practice), summarize,
              mean_LR=mean(mean_LR))

```

### Proportion of Income Due to Labor Return

```{r, fig.pos='h', fig.width=9, fig.align='center',echo=FALSE}
qplot(data=lr_effect_crop, x=year, y=prop_LR, colour=crop, group=crop, geom="line") + facet_wrap(~practice) + theme(legend.position="bottom")

qplot(data=lr_effect, x=year, y=prop_LR, colour=practice, group=practice, geom="line") + theme(legend.position="bottom")
```

The plots above show what proportion of Thompson and Boone total income is accounted for by Labor Return.\\

Thompson farms have higher average labor returns in general, but this shows that the labor return appears to account for a higher proportion of income in the Thompson farms than in the Boone farms. Do the work yourself, if you can! 

### Proportion of Thompson Advantage Due to Labor Return
```{r, fig.pos='h', fig.width=7, fig.align='center',echo=FALSE}
diff_profit<-subset(total_income, practice=="Thompson")$mean_income - subset(total_income, practice=="Boone")$mean_income
diff_lr<-subset(lr, practice=="Thompson")$mean_LR - subset(lr, practice=="Boone")$mean_LR

df2<-data.frame(year=unique(total_income$year), diff_profit, diff_lr, prop_lr=diff_lr/diff_profit)
qplot(data=df2, x=year, y=prop_lr, geom="line", xlab="Year", ylab="Proportion of Profit Difference") 
```

Consistently, between 10-20% of Thompson's advantage is due to the the difference in labor return between Boone and Thompson farms.

## Effect of Livestock

```{r, echo=FALSE, warning=FALSE}
### Income and Expenses ###
income_crop<-ddply(years2, .(year, practice, crop), summarize,
                   mean_income = mean(Crop_Income))
income<-ddply(income_crop, .(year, practice), summarize,
              mean_income=mean(mean_income))

expense_crop<-ddply(dat2[dat2$item_type=="Expense",], .(year, field, crop), summarize,
                    expense = sum(value))
expense_crop$practice<-as.factor(sub("[0-9].*", "", expense_crop$field))

### Remove fertilizer
expense_nofert_crop<-ddply(dat2[dat2$item_type=="Expense" & dat2$item!="Purch_Pert" & dat2$item!="Maunure_Charge",], .(year, field, crop), summarize,
                    expense = sum(value))

expense_nofert_crop$practice<-as.factor(sub("[0-9].*", "", expense_crop$field))

expense_crop2<-ddply(expense_crop, .(year, practice, crop), summarize,
                     expense=mean(expense))
expense_nofert_crop2<-ddply(expense_nofert_crop, .(year, practice, crop), summarize,
                     expense_nofert=mean(expense))

expenses<-merge(expense_crop2, expense_nofert_crop2)

expenses_overall<-ddply(expenses, .(year, practice), summarize,
               mean_expense=mean(expense),
               mean_expense_nofert=mean(expense_nofert))

add_income<-ddply(years2, .(year, practice, crop), summarize,
                extra_income=mean(Pasture_per_Stubble + Residue_Income + Straw_Income))

income_crop$extra<-add_income$extra_income
income2<-ddply(income_crop, .(year, practice), summarize,
               mean_allincome=mean(mean_income+extra))

df<-merge(income, income2)
df<-merge(df, expenses_overall)

df$rawprofit <- df$mean_income - df$mean_expense
df$profit_extra <- income2$mean_allincome - df$mean_expense
df$rawprofit_nofert <- df$mean_income - df$mean_expense_nofert

```


### With/Without Fertilizer Costs

The profit and losses from livestock on the Thompson farms are not included in the data. It is suggested that Thompson fertilized his fields using manure from his livestock, and subsequently has an expense due to spreading manure, but no expenses for purchasing fertilizer. The Boone expenses on the other hand include a large sum (relative to other expenses) for purchasing fertilizer. As livestock costs and profits are not included, it seems reasonable to explore excluding these variables from the analysis. Note: this analysis was done without additional income from straw, stubble, and residue.

#### Expenses
```{r, fig.pos='h', fig.width=9, fig.align='center',echo=FALSE}
theme_set(theme_bw(base_size=12))
p1<-qplot(data=df, x=year, y=mean_expense, group=practice, colour=practice, geom="line", ylab="Expense/Acre", xlab="Year", main="Average Expenses by Farm") + theme(legend.position="bottom")
p2<-qplot(data=df, x=year, y=mean_expense_nofert, group=practice, colour=practice, geom="line", ylab="Expense/Acre", xlab="Year", main="Average Expenses w/o Fertilizer by Farm") + theme(legend.position="bottom")
grid.arrange(p1,p2,ncol=2)
```

#### Profit 
```{r, fig.pos='h', fig.width=9, fig.align='center',echo=FALSE}
p5<-qplot(data=df, x=year, y=rawprofit, group=practice, colour=practice, geom="line", ylab="Profit/Acre", xlab="Year", main="Average Profit") + theme(legend.position="bottom")
p6<-qplot(data=df, x=year, y=rawprofit_nofert, group=practice, colour=practice, geom="line", ylab="Profit/Acre", xlab="Year", main="Average Profit w/o Fertilizer") + theme(legend.position="bottom")
grid.arrange(p5,p6,ncol=2)
```

Removal of fertilizer affects seems to lessen the difference in expenses and profit between the Boone and Thompson averages, but the Thompson farm still consistently comes out ahead. 

#### With/Without Fertilizer Costs -- Just Corn vs SB

```{r, fig.pos='h', fig.width=8, fig.height=8, fig.align='center',echo=FALSE}
p3<-qplot(data=subset(expenses, crop=="Corn" | crop=="SB"), x=year, y=expense, group=practice, colour=practice, geom="line", ylab="Expense/Acre", xlab="Year", main="Average Expenses") + facet_wrap(~crop) + theme(legend.position="bottom")
p4<-qplot(data=subset(expenses, crop=="Corn" | crop=="SB"), x=year, y=expense_nofert, group=practice, colour=practice, geom="line", ylab="Expense/Acre", xlab="Year", main="Average Expenses w/o Fertilizer") + facet_wrap(~crop) + theme(legend.position="bottom")
grid.arrange(p3,p4,ncol=1)
```

When just looking at corn, expenses for Boone drop considerably and decrease the gap between farms.

```{r, echo=FALSE, warning=FALSE}
### Remove fertilizer
expense_nofert_crop<-ddply(subset(dat2, item_type=="Expense" & item!="Purch_Pert" & item!="Maunure_Charge" & item!="Windrow_Oats" & item!="Stubble_Costs" & item!="Straw_Costs"), .(year, field, crop), summarize,
                    expense = sum(value))

expense_nofert_crop$practice<-as.factor(sub("[0-9].*", "", expense_crop$field))

expense_nofert_crop2<-ddply(expense_nofert_crop, .(year, practice, crop), summarize,
                     expense_nofert=mean(expense))

expenses<-merge(expense_crop2, expense_nofert_crop2)

expenses_overall<-ddply(expenses, .(year, practice), summarize,
               mean_expense=mean(expense),
               mean_expense_nofert=mean(expense_nofert))

add_income<-ddply(years2, .(year, practice, crop), summarize,
                extra_income=mean(Pasture_per_Stubble + Residue_Income + Straw_Income))

income_crop$extra<-add_income$extra_income
income2<-ddply(income_crop, .(year, practice), summarize,
               mean_allincome=mean(mean_income+extra))

df<-merge(income, income2)
df<-merge(df, expenses_overall)

df$rawprofit <- df$mean_income - df$mean_expense
df$profit_extra <- income2$mean_allincome - df$mean_expense
df$rawprofit_nofert <- df$mean_income - df$mean_expense_nofert
```

### All Additional Income and Expenses
The exploratory analysis aboves focuses on crop income, ignoring additional income from straw, stubble, and residue. However, we should also remove costs related with these income sources if we don't include them (Purch\_Pert, Manure\_Charge, Windrow\_Oats, Stubble\_Costs, Straw\_Costs).

#### Expenses
```{r, fig.pos='h', fig.width=9, fig.align='center',echo=FALSE}
p1<-qplot(data=df, x=year, y=mean_expense, group=practice, colour=practice, geom="line", ylab="Expense/Acre", xlab="Year", main="Average Expenses by Farm") + theme(legend.position="bottom")
p2<-qplot(data=df, x=year, y=mean_expense_nofert, group=practice, colour=practice, geom="line", ylab="Expense/Acre", xlab="Year", main="Average Expenses w/o Additional Practices") + theme(legend.position="bottom")
grid.arrange(p1,p2,ncol=2)
```

#### Profit
```{r, fig.pos='h', fig.width=9, fig.align='center',echo=FALSE}

p5<-qplot(data=df, x=year, y=profit_extra, group=practice, colour=practice, geom="line", ylab="Profit/Acre", xlab="Year", main="Average Profit") + theme(legend.position="bottom")
p6<-qplot(data=df, x=year, y=rawprofit_nofert, group=practice, colour=practice, geom="line", ylab="Profit/Acre", xlab="Year", main="Average Profit w/o Additional Practices") + theme(legend.position="bottom")
grid.arrange(p5,p6,ncol=2)
```

```{r, fig.pos='h', fig.width=6, fig.align='center',echo=FALSE}
diff_pextra<-subset(df, practice=="Thompson")$profit_extra - subset(df, practice=="Boone")$profit_extra
diff_pnone<-subset(df, practice=="Thompson")$rawprofit_nofert - subset(df, practice=="Boone")$rawprofit_nofert
df2<-data.frame(year=rep(unique(df$year),2), diff<-c(diff_pextra, diff_pnone), method=rep(c("Addition Practices", "No Additional Practices"), each=length(diff_pextra)))

qplot(data=df2, x=year, y=diff, group=method, colour=method, geom="line", ylab="Difference in Profit", xlab="Year", main="Thompson Advantage") + geom_hline(aes(yintercept=0)) + theme(legend.position="bottom")
```

After accounting for these additional variables, we still see that the differences in profit between Thompson's farm and the Boone averages decreases. Thompson still has consistently higher profits across the years, but it is interesting to see that a ``large" portion of the advantage was due to these fertilizer effects.