---
title: "Expenses"
author: "Colin"
date: "Saturday, October 25, 2014"
output: html_document
---

#### Libraries 
```{r,warning=FALSE,message=FALSE}
library(ggplot2)
library(dplyr)

##### Read in Data #####
pfi <- read.csv("C:/Users/Colin LB/Documents/GitHub/PFI/data/PFI_clean.csv")
#Let's remove Land_Change since that's assumed the same for both farms.
##### Exploring #####
pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(crop %in% c("Corn")) %>%
  filter(item !="Land_Change") %>%
  filter(value > .01) %>%
  group_by(farmer, year, crop,field_id,item) %>%
  summarise(total = sum(value)) %>%
  group_by(farmer, year,item) %>%
  summarise(avg_total = mean(total)) %>%
  ggplot() +
  geom_line(aes(x=year, y=avg_total, colour=item)) +
  facet_wrap(~farmer) +
  ggtitle('Expenses over time for Corn Averaged Over Plots by Item')
```

####Let's just pull out the costs that seem "large"

```{r}
pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(crop %in% c("Corn")) %>%
  filter(item %in% c("Purch_Pert","Seed","Herbicides")) %>%
  group_by(farmer, year, crop, field_id,item) %>%
  summarise(total = sum(value)) %>%
  group_by(farmer,year,item) %>%
  summarise(avg_total = mean(total)) %>%
  ggplot() +
  geom_line(aes(x=year, y=avg_total,color=item)) +
  facet_wrap(~farmer) +
  ggtitle('Expenses over time for Corn Averaged over Plots by Item')
```

Did we confirm what Purch_Pert is?  Let's look just at SB now.
```{r}
pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(crop %in% c("SB")) %>%
  filter(item !="Land_Change") %>%
  filter(value > .01) %>%
  group_by(farmer, year, crop,field_id,item) %>%
  summarise(total = sum(value)) %>%
  group_by(farmer, year,item) %>%
  summarise(avg_total = mean(total)) %>%
  ggplot() +
  geom_line(aes(x=year, y=avg_total, colour=item)) +
  facet_wrap(~farmer) +
  ggtitle('Expenses over time for SB Averaged Over Plots by Item')
```

####Let's look at the "large" expenses for SB

```{r}
pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(crop %in% c("SB")) %>%
  filter(item %in% c("Purch_Pert","Seed","Herbicides")) %>%
  group_by(farmer, year, crop, field_id,item) %>%
  summarise(total = sum(value)) %>%
  group_by(farmer,year,item) %>%
  ggplot() +
  geom_line(aes(x=year, y=total,color=item)) +
  facet_wrap(~farmer) +
  ggtitle('Expenses over time for SB')
```


####Let's try and see all the expenses Thompson records as 0 vs. what Boone pays.


q<-pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(crop %in% c("Corn")) %>%
  filter(item !="Land_Change") %>%
  filter(value==0 &farmer=="Thompson",field_id==1) %>%
  group_by(item)
  
unique(q$item)


```{r}
boone_extra_expense<-pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(farmer=="Boone") %>%
  filter(crop %in% c("Corn")) %>%
  filter(item %in% c("Apply_NH4","Hedge_per_PL","Corn_RSL","Herbicides","Spring_Tillage","Chop_StksCc","Spray_per_Walk","Purch_pert","Fall_Tillage")) %>%
  filter(value>.01) %>%
  group_by(year,item) %>%
  summarise(total = sum(value)) %>%
  ggplot() +
  geom_line(aes(x=year, y=total,linetype=item)) +
  ggtitle('Boone Extra Expenses for Corn')

boone_extra_expense
```


####Hedge Per PL and Herbicide Are Again the big Difference



