---
title: "Expenses"
author: "Colin"
date: "Saturday, October 25, 2014"
output: html_document
---

##### Libraries #####
library(ggplot2)
library(dplyr)
library(reshape2)

##### Read in Data #####
pfi <- read.csv("C:/Users/Colin LB/Documents/GitHub/PFI/data/PFI_clean.csv")

##### Exploring #####
```{r}
z2<-pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(crop %in% c("Corn")) %>%
  group_by(farmer, year, crop, field_id,item) %>%
  summarise(total = sum(value)) %>%
  group_by(farmer, year, crop,item) %>%
  ggplot() +
  geom_line(aes(x=year, y=total, colour=item)) +
  facet_wrap(~farmer) +
  ggtitle('Expenses over time for Corn')
```
Let's just pull out the costs that seem "large"

```{r}
q<-pfi[complete.cases(pfi),] %>% 
  filter(item_type == "Expense") %>%
  filter(crop %in% c("Corn")) %>%
  filter(item %in% c("Land_Change","Purch_Pert","Seed","Herbicides")) %>%
  group_by(farmer, year, crop, field_id,item) %>%
  summarise(total = sum(value)) %>%
  group_by(farmer,year,item) %>%
  ggplot() +
  geom_line(aes(x=year, y=total,color=item)) +
  facet_wrap(~farmer) +
  ggtitle('Expenses over time for Corn')
```


