---
title: 'PFI: Yields'
author: "Evan “Pete” Walsh"
date: "October 27, 2014"
output: pdf_document
---

\subsection*{Yields for SB and Corn in acre buschel per pound.}

Read data and load libraries.

```{r message=FALSE}
pfi <- read.csv("/Users/marianwaitwalsh/GitHub/PFI/data/PFI_clean.csv")
weather <- read.csv("/Users/marianwaitwalsh/GitHub/PFI/data/IA_annual_rainfall_raw.csv")
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(leaps)
```

Subset the ```PFI``` data to get yields for just corn and SB.

```{r}
yields <- pfi %>%
  filter(item_type == "Unit Quantity", crop %in% c("Corn","SB")) %>%
  select(-c(item, item_type)) %>%
  group_by(field_id, year)
```

Clean the weather data and join with ```yields```.

```{r warning=FALSE}
wBoone <- weather %>%
  filter(stationName == "Boone") %>% 
  gather(key, value, 5:373) %>%
  separate(key, into = c("year", "key"), sep = "\\_") %>%
  spread(key, value) %>%
  select(year, MAXT, MINT, PREC)

wBoone$year <- sapply(wBoone$year, FUN = function(x) extract_numeric(x))
yields2 <- inner_join(yields, wBoone, by = "year")
yields2$MAXT <- as.numeric(yields2$MAXT)
yields2$MINT <- as.numeric(yields2$MINT)
yields2$PREC <- as.numeric(yields2$PREC)
```

Examine the relationship between annual precipitation ```PREC```, mean maximum temperature ```MAXT``` and mean minimum temperature ```MINT```.

```{r fig.height=4, fig.width=7}
head(yields2)
qplot(PREC, value, data=yields2, colour=farmer, shape=farmer,
      facets=~crop) + geom_smooth(method="lm", se=F) +
  ylab("Yield")
qplot(MAXT, value, data=yields2, colour=farmer, shape=farmer,
      facets=~crop) + geom_smooth(method="lm", se=F)+
  ylab("Yield")
qplot(MINT, value, data=yields2, colour=farmer, shape=farmer,
      facets=~crop) + geom_smooth(method="lm", se=F)+
  ylab("Yield")
```

Overall, pretty weak relationships between ```Yield``` for ```Corn``` and ```SB``` and ```PREC```, ```MAXT```, and ```MINT```. Perhaps we need to look at monthly or seasonal weather data instead to see more of a trend.

Fields by year for Thompson that were used for ```SB``` or ```corn```.

```{r}
# pfi %>% 
#   filter(item_type == "Unit Quantity", crop %in% c("Corn","SB"),
#          farmer == "Thompson") %>%
#   group_by(year, field_id) %>%
#   select(1:4)
```

Examine how farming practices relate to ```Yield``` by turning the ```Expense``` data into indicator variables:
\begin{itemize}
\item Each expense category is an indicator variable
\item Variable is marked \textbf{TRUE} if the expense for that category for each particular \textbf{field.id} and \textbf{year} is greater than 0.
\item \textbf{FALSE} if 0.
\end{itemize}

```{r}
yields3 <- pfi %>%
  filter(item_type == "Expense", crop  %in% c("Corn", "SB")) %>% 
  spread(item, value) %>% 
  select(-c(5))
for (i in 5:36) {
  yields3[,i] <- yields3[,i] != 0
}
yields3$yield <- yields$value
yields3 <- yields3[,c(1:4,37,5:36)] # move 'yield' to 5th column
head(yields3)[1:6]
```

Eliminate variables that are all ```TRUE``` or all ```FALSE```.

```{r}
N <- nrow(yields3)
C <- ncol(yields3)
idx <- NULL
for (i in 6:C) {
  if ((sum(yields3[,i]) == N) | (sum(yields3[,i] == FALSE) == N)) {
    idx <- c(idx, i)
  }
}
yields3 <- yields3[-idx]
names(yields3)
```

10 indicator variables eliminated so far. Break the dataset into yields for ```Corn``` and yields for ```SB```. Elimate variables again that are all ```TRUE``` or all ```FALSE```.

```{r}
yields3_C <- subset(yields3, crop == "Corn")
yields3_SB <- subset(yields3, crop == "SB")

# yields3_C doesn't have any variables that are all T or all F

N <- nrow(yields3_SB)
C <- ncol(yields3_SB)
idx <- NULL
for (i in 6:C) {
  if ((sum(yields3_SB[,i]) == N) | (sum(yields3_SB[,i] == FALSE) == N)) {
    idx <- c(idx, i)
  }
}
yields3_SB <- yields3_SB[-idx]
```

Run regsubets to find best model for ```Corn``` yields.

```{r warning=FALSE, message=FALSE}
regsubsets.out <- regsubsets(x = as.matrix(yields3_C[,6:27]), 
                             y = yields3_C[,5])
# Variables with black boxes at the highest y-axis label should be included
plot(regsubsets.out, scale = "adjr2", main = "Adjusted R^2 Corn Model")
summary(lm(data=yields3_C, yield~Apply_NH4+Crop_Ins+Interest+
             Mov_and_Stor_bales+Mow_per_Windrow+Rake+Rotary_Hoe+
             Spray_per_Walk+Stack_Residues))
```

Note that when NH4 is applied, the expected yield decreases by 51.892 when everything else is held constant. Why is crop insurance such a good predictor? Maybe the farmer buys crop insurance when they think it's going to be a bad year. Or it's just because Thompson stopped using crop insurance after 2002, and has had consistently higher yields than the Boone average.

Here's a model that makes more "sense":

```{r}
summary(lm(data=yields3_C, yield~Apply_NH4+
             Mow_per_Windrow+Rotary_Hoe+Spray_per_Walk))
```

Run regsubsets for best model for ```SB``` yields.

```{r warning=FALSE}
regsubsets.out2 <- regsubsets(x = as.matrix(yields3_SB[,6:22]), 
                             y = yields3_SB[,5])
plot(regsubsets.out2, scale = "adjr2", main = "Adjusted R^2 SB Model")
summary(lm(data=yields3_SB, yield~Chop_StksCc+Cover_Crop+Crop_Ins+
             Cultivation+Maunure_Charge+Mov_and_Stor_bales+Rotary_Hoe+
             Stack_Residues))
```

Crop insurance again is found to be a significant predictor.

Here is a model that makes more "sense":

```{r}
summary(lm(data=yields3_SB, yield~Cultivation+Maunure_Charge+Rotary_Hoe))
```

